Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A cross-sectional study based on the parametric g-formula

### 2024-07-19 output/paper/_si.qmd
Finished addressing comments reviwer 1.

m1sa$df_gt
```
**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::

@@

m2sa$df_gt
```
**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
:::

@@

m3sa$df_gt
```
**Marginal contrasts for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::

>>>>

m1sa$table
```
**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::

@@

m2sa$table
```
**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the {{< acr edcs first_use=true >}} on the glucocorticosteroids expressed in ng/ml (HELIX subcohort; 2013-2016).**
:::

@@

m3sa$table
```
**Marginal contrasts on the logarithmic scale for effect modification by sex of a increase from the 10th to the 90th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::

### 2024-07-19 output/paper/manuscript.qmd
Finished addressing comments reviwer 1.

@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolite MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)). The individual {{< acr mcs >}} are presented in Supplementary Tables [-@supptbl-marginal-1sa], [-@supptbl-marginal-2sa] and [-@supptbl-marginal-3sa], and in Supplementary Figures [-@suppfig-marginal-1sa], [-@suppfig-marginal-2sa] and [-@suppfig-marginal-3sa].

@Tbl-hypothesis-3 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}. Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost. prod.", "HRT-SE")`). Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females). The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-3sa.

@@

We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain [@LupienMcEwenGunnar:2009].

Our findings should be interpreted in light of the following strengths and limitations. Strengths include its relative large sample size and its inclusion of multiple classes of non-persistent {{< acr edcs >}}. Further, this study used pooled urine samples for chemical assessment to obtain more representative long-term exposures, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018]. We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to obtain *doubly robust* effect estimates. Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure.

>>>>

@Tbl-hypothesis-1and2 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}.
For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolite MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`).
For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the phenol ETPA (corticosterone production, `r format_res(heffects2, "etpa", "corticosterone production")`) and the phthalate metabolite MEHP (cortisol production, (`r format_res(heffects2, "mehp", "cortisol production")`); cortisone production, (`r format_res(heffects2, "mehp", "cortisone production")`)). The individual {{< acr mcs >}} are presented in Supplementary Tables [-@supptbl-marginal-1sa] and [-@supptbl-marginal-2sa], and in Supplementary Figures [-@suppfig-marginal-1sa] and [-@suppfig-marginal-2sa].

@Tbl-hypothesis-3 presents the results of the difference between estimates of the {{< acr mcs >}} on the logarithmic scale for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}. Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost. prod.", "HRT-SE")`). Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females). The individual {{< acr mcs >}} are presented in @supptbl-marginal-3sa and in @suppfig-marginal-3sa.

@@

We are also not aware of prior epidemiological studies specifically investigating the effects of elevated levels of glucocorticosteroids in relation to attentional function, although there is evidence that under- or over-production of glucocorticosteroids interfere with the normal development of the brain, possibly in a sex-specific manner [@LupienMcEwenGunnar:2009].

Our findings should be interpreted in light of the following strengths and limitations. Strengths include its relative large sample size and its inclusion of multiple classes of non-persistent {{< acr edcs >}}. Further, this study used pooled urine samples for chemical assessment to reduce measurement error, since it is known that these specific {{< acr edcs >}} have very short half-lives [@PerrierGiorgis-AllemandSlama:2016; @CasasBasaganaSakhi:2018]. We decided to model both the *treatment* mechanisms, for the estimation of balancing weights, and the outcomes, with traditional covariates adjustment, to obtain *doubly robust* effect estimates. Finally, we decided not to interpret our results by focusing on the estimated coefficients of possibly misspecified regression models, but by making use of the g-computation procedure.
